Type Evolution in Programming Systems

### 2024-05-24 paper.tex
3D divergence diagram

\usepackage{csquotes}
\usepackage{changepage}
\usepackage{multicol}
\usepackage{ccicons}
\usepackage[many]{tcolorbox}
\lstdefinelanguage[programming]{TeX}[AlLaTeX]{TeX}{%
  deletetexcs={title,author,bibliography},%
  deletekeywords={tabular},
  morekeywords={abstract},%
  moretexcs={chapter},%
  moretexcs=[2]{title,author,subtitle,keywords,maketitle,titlerunning,authorinfo,affiliation,authorrunning,paperdetails,acks,email},
  moretexcs=[3]{addbibresource,printbibliography,bibliography},%
}%
\lstset{%
  language={[programming]TeX},%
  keywordstyle=\firamedium,
  stringstyle=\color{RosyBrown},%
  texcsstyle=*{\color{Purple}\mdseries},%
  texcsstyle=*[2]{\color{Blue1}},%
  texcsstyle=*[3]{\color{ForestGreen}},%
  commentstyle={\color{FireBrick}},%
  escapechar=`,}
\newcommand*{\CTAN}[1]{\href{http://ctan.org/tex-archive/#1}{\nolinkurl{CTAN:#1}}}
%%

@@

>>>>

\usepackage{csquotes}
\usepackage{changepage}
\usepackage{multicol}
\usepackage{ccicons}
\usepackage[many]{tcolorbox}
\usepackage{tikz-cd}
\usepackage{float}

\lstdefinelanguage[programming]{TeX}[AlLaTeX]{TeX}{%
  deletetexcs={title,author,bibliography},%
  deletekeywords={tabular},
  morekeywords={abstract},%
  moretexcs={chapter},%
  moretexcs=[2]{title,author,subtitle,keywords,maketitle,titlerunning,authorinfo,affiliation,authorrunning,paperdetails,acks,email},
  moretexcs=[3]{addbibresource,printbibliography,bibliography},%
}%
\lstset{%
  language={[programming]TeX},%
  keywordstyle=\firamedium,
  stringstyle=\color{RosyBrown},%
  texcsstyle=*{\color{Purple}\mdseries},%
  texcsstyle=*[2]{\color{Blue1}},%
  texcsstyle=*[3]{\color{ForestGreen}},%
  commentstyle={\color{FireBrick}},%
  escapechar=`,}
\newcommand*{\CTAN}[1]{\href{http://ctan.org/tex-archive/#1}{\nolinkurl{CTAN:#1}}}
%%

@@

\section{diagrams}

\begin{figure}[H]
  \begin{tikzcd}[sep = 6em]
    \textsf{query}
    \arrow[d, "\textsf{refactor}_\alpha"{name=L, description}]
    &
    \textsf{schema}
    \arrow[l, "\textsf{constrains}"{above}]
    \arrow[r, "\textsf{constrains}"{above}]
    \arrow[d, red, "\textsf{change}_\alpha"{name=C, description}]
    &
    \textsf{data}
    \arrow[d, "\textsf{migrate}_\alpha"{name=R, description}]
    \\
    \textsf{query}_\alpha
    &
    \textsf{schema}_\alpha
    \arrow[l, "\textsf{constrains}"{above}]
    \arrow[r, "\textsf{constrains}"{above}]
    &
    \textsf{data}_\alpha
    \arrow[Rightarrow, from=C, to=L, "\textsf{propagate}"{above}]
    \arrow[Rightarrow, from=C, to=R, "\textsf{propagate}"{above}]
  \end{tikzcd}
  \caption{Extract Entity}
\end{figure}

\begin{figure}[H]
  \begin{tikzcd}[column sep = large, row sep = 5em]
    &
    \textsf{schema}
    \arrow[ld, red, "\textsf{change}_\alpha"{name=L2, near start, description}]
    \arrow[rr, "\textsf{constrains}"]
    \arrow[dd, red, "\textsf{change}_\beta"{name=L, near start, description}]
    &
    &
    \textsf{data}
    \arrow[dd, "\textsf{migrate}_\beta"{name=R, near start, description}]
    \arrow[Rightarrow, from=L, to=R]
    \\
    \textsf{schema}_\alpha
    %\arrow[dd, "\textsf{change}_\beta"{name=L3, near start, description}]
    \arrow[rr, crossing over]
    &
    &
    \textsf{data}_\alpha
    \arrow[from=ur, crossing over, "\textsf{migrate}_\alpha"{name=R2, near start, description}]
    \arrow[Rightarrow, from=L2, to=R2, crossing over, "\textsf{propagate}"{above}]
    &
    \\
    &
    \textsf{schema}_\beta
    \arrow[ld, red, "\mathit{merge}\left(\textsf{change}_\alpha {\,,\;} \textsf{change}_\beta\right)"{name=L4, near start, description}]
    \arrow[rr]
    &
    &
    \textsf{data}_\beta
    \arrow[ld, "\textsf{migrate}_{\alpha\beta}"{name=R4, near start, description}]
    \arrow[Rightarrow, from=L4, to=R4]
    \\
    \textsf{schema}_{\alpha\beta}
    \arrow[rr]
    &
    &
    \textsf{data}_{\alpha\beta}
    %\arrow[from=uu, crossing over, "\textsf{migrate}_\beta"{name=R3, near start, description}]
    %\arrow[Rightarrow, from=L3, to=R3, crossing over]
    &
  \end{tikzcd}
  \caption{Mergeing divergent changes}
\end{figure}
